Determinism, Complexity, and Predictability in Computer Performance

### 2014-02-10 pecomplexity.tex
adding outines

\section{Introduction}
%\begin{it}
%Paragraph on computer performance, including citations to Todd paper
%and summary of the results that indicate that they're deterministic
%nonlinear dynamical systems.  Given that, we should be able to
%predict.  What benefits would accrue if we could do so: power mgmt,
%end world hunger [[this is my primary goal everyday :)]], etc.
%\end{it}

@@

\begin{enumerate}
\item \cmark~Introduce MASE.
\item \cmark~WPE is a good measure of predictability.  Figure: best athlete MASE vs.
its WPE.
\item \cmark~Talk about structure analysis. Just because there is forward information
transfer, does not mean that linear predictors can get at this.
\item For this show a figure of (a) ARIMA vs MASE (b) LMA vs MASE in a side by
side plot.
\item full results.  Image: MASE vs. WPE for both LMA \& ARIMA.  Points to make:
\begin{enumerate}
\item \cmark~clusters are distributed differently
\item \cmark~clusters are shaped differently---tight or not
\item \cmark~clusters move differently between LMA and ARIMA
\item finally, the diagonal line is important. If you're below it, you could do
better.
\end{enumerate}
\end{enumerate}

>>>>

\section{Introduction}
%\begin{it}
%Paragraph on computer performance, including citations to Todd paper
%and summary of the results that indicate that they're deterministic
%nonlinear dynamical systems.  Given that, we should be able to
%predict.  What benefits would accrue if we could do so: power mgmt,
%end world hunger [[this is my primary goal everyday :)]], etc.
%\end{it}
Things to add to introduction
\begin{enumerate}
\item Differnt kinds of complexity exist in time series and this makes choosing prediction models difficult 
\item When deterministic structure even complex structure exists that structure can be utilized for prediction. 
\item For noisy real-valued time series distinguishing randomness (WN,RW) complexity from structured nonlinear / chaotic /high period / high dimensional etc complexity is often impossible.

\item Make an argument that Computer Performance is a great testing ground as it omits signals that completely cover the spectrum of complexity \col \dots \gcc

\item Introduce the two main contributions of the paper which are outlined at the begining of the results section

@@

OUTLINE:
\begin{enumerate}
\item \cmark~Introduce MASE.
\item \cmark~WPE is a good measure of predictability.  Figure: best athlete MASE vs.
its WPE.
\item \cmark~Talk about structure analysis. Just because there is forward information
transfer, does not mean that linear predictors can get at this.
\subitem \cmark~For this show a figure of (a) ARIMA vs MASE (b) LMA vs MASE in a side by
side plot.
\item full results.  Image: MASE vs. WPE for both LMA \& ARIMA.  Points to make:
\begin{enumerate}
\item \cmark~clusters are distributed differently
\item \cmark~clusters are shaped differently---tight or not
\item \cmark~clusters move differently between LMA and ARIMA
\item finally, the diagonal line is important. If you're below it, you could do
better.
\end{enumerate}
\end{enumerate}

In this section we will validate the two key findings of this work:

[[Rephrase these]]
\begin{enumerate}

\item The existence of predictable structure in noisy real-valued time series is quantifiable by WPE and as a result WPE is correlated with prediction accuracy (MASE)

\subitem WPE can quantify when a noisy real-valued time series is predictable.[[I am unsatisfied with predictable in this sentence, need a better word to say "better than random walk" or ``able to be forecast effectively" or ``has the structural capacity to transmit information in a way that the time series can be effectively forecast"

%\item The existence of complex structure in a noisy real-valued time series is quantifiable and this type of complexity is directly correlated with predictability

\item The way structure/information/complexity is processed internally by a given process plays a crucial role in predictability.
\subitem We will have shown that the existence whether linear or nonlinear is picked up on with WPE but this point gets at whether the prediction model can use the structure or not (linear can't use nonlinear structure). The right to left shifts in \col and some of the \svd regimes and the lack of shift in \gcc illustrate this nicely.

\subitem This is the linear vs nonliear vs random we see with the right to left shifts with lower complexity time series.
